|

### 2024-08-20 reproducibility/manuscript/manuscript.qmd
docs(manuscript): lint
Signed-off-by: Cameron Smith <cameron.ray.smith@gmail.com>

---
title: "Pyro&thinsp;-Velocity: Probabilistic RNA Velocity inference from single-cell data"
number-sections: true
toc: true
toc-title: Contents
format:
  nature-pdf:
    # journal.cite-style is included in the tex file but ignored by pandoc if 
    # cite-method is not `natbib`.
    journal:
      cite-style: sn-nature
    classoption: 
      - lineno
    # `citeproc` is the pandoc default. Set `cite-method: natbib` if required
    # to use the bst styles from the upstream template.
    cite-method: natbib
    keep-tex: true
    equal-margins: false
    include-in-header: header.tex
    link-citations: true
  html:
    code-fold: true
    html-math-method: mathjax
  docx:
    reference-doc: styles.docx
# filters:
#   - authors-block
csl: _extensions/sciexp/nature/csl/springer-basic-brackets-no-et-al.csl
affiliations:
  - name: Massachusetts General Hospital Research Institute, Charlestown, MA
    id: 1
    department: Molecular Pathology Unit
    city: Charlestown
    state: MA
    postal-code: 02129
  - name: Harvard Medical School, Charlestown, MA
    id: 2
    department: Massachusetts General Hospital Cancer Center
    city: Charlestown
    state: MA
    postal-code: 02129
  - name: Massachusetts General Hospital, Boston, MA
    id: 3
    department: Center for Regenerative Medicine
    city: Boston
    state: MA
    postal-code: 02114
  - name: Harvard University, Cambridge, MA
    id: 4
    department: Harvard Stem Cell Institute
    city: Cambridge
    state: MA
    postal-code: 02139
  - name: Cambridge, MA
    id: 5
    department: Broad Institute of MIT and Harvard
    city: Cambridge
    state: MA
    postal-code: 02142
  - name: Laboratory of Neurodevelopmental Systems Biology, Lausanne, Switzerland
    id: 6
    department: Brain and Mind Institute, School of Life Sciences, École Polytechnique Fédérale de Lausanne (EPFL)
    city: Lausanne
    state: Switzerland
    postal-code: 1015
  - name: Cambridge, MA
    id: 7
    department: Basis Research Institute
    city: Cambridge
    state: MA
    postal-code: 02142
authors:
  - name: Qian Qin
    affiliations:
      - ref: 1
      - ref: 2
      - ref: 3
      - ref: 4
      - ref: 5
    # attributes:
    #   equal-contributor: true
  - name: Eli Bingham
    affiliations:
      - ref: 5
      - ref: 7
  - name: Gioele La Manno
    affiliations:
      - ref: 6
    email: gioele.lamanno@epfl.ch
    attributes:
      corresponding: true
  - name: David M. Langenau
    affiliations:
      - ref: 1
      - ref: 2
      - ref: 3
      - ref: 4
    email: dlangenau@mgh.harvard.edu
    attributes:
      corresponding: true
  - name: Luca Pinello
    affiliations:
      - ref: 1
      - ref: 2
      - ref: 5
    email: lpinello@mgh.harvard.edu
    attributes:
      corresponding: true
abstract: |
  Single-cell RNA Velocity has dramatically advanced our ability to model cellular differentiation and cell fate decisions. However, current preprocessing choices and model assumptions often lead to errors in assigning developmental trajectories. Here, we develop, Pyro-Velocity, a Bayesian, generative, and multivariate RNA Velocity model to estimate the uncertainty of cell future states. This approach models raw sequencing counts with the synchronized cell time across all expressed genes to provide quantifiable and improved information on cell fate choices and developmental trajectory dynamics.
keywords: [RNA velocity, single-cell RNA sequencing, probabilistic generative modeling, cell fate, developmental trajectory inference]
bibliography: bibliography.bib
crossref:
  fig-prefix: Fig.
  custom:
    - kind: float
      key: suppfig
      latex-env: suppfig
      reference-prefix: Fig. S
      space-before-numbering: false
      latex-list-of-description: Supplementary Figure
---

@@

# Introduction {#sec-intro}

The Introduction section, of referenced text [@greenwade93] expands on the background of the work (some overlap with the Abstract is acceptable). The introduction should not include subheadings.

Springer Nature does not impose a strict layout as standard however authors are advised to check the individual requirements for the journal they are planning to submit to as there may be journal-level preferences. When preparing your text please also be aware that some stylistic choices are not supported in full text XML (publication version), including coloured font. These will not be replicated in the typeset article if it is accepted.

# Using this template {#sec-template}

This Quarto template is unofficial and built out of Springer Nature's template. 
Some examples of commonly used commands and features are listed below, to help you get started.

As seen below, you can mix markdown and Latex with each other, though it's likely best to mostly use markdown.

## Cross Referencing

To reference a figure with example label "plot", use `@fig-plot`. 
Analogously, to reference a table with example label "data", use `@tbl-data`.
To reference a section, such as the Introduction (@sec-intro), use `@sec-intro`.

For complete information on cross referencing with Quarto, see <https://quarto.org/docs/authoring/cross-references.html>.

## Citations

Quarto formats citations and references automatically using the bibliography records in your .bib file. For a single citation use [@greenwade93]. For multiple citations use [@greenwade93;@knuth1984texbook]. If using an author-date-based citation style, then for a citation in parentheses use `[@greenwade93]` and for a text citation use `@greenwade93`. You can drop the author using `[-@greenwade93]`. If the tex output is to be included in a submission to a preprint server or publisher, the default citation method, `citeproc`, may not always produce bibliographies compatible with the provided `bst` styles. In this case, you can set `cite-method: natbib`. See <https://quarto.org/docs/authoring/footnotes-and-citations.html#sec-biblatex>. In this case, you may either restrict usage to the default bracketed key citation formats supported by pandoc, such as [@greenwade93] or [@greenwade93;@knuth1984texbook]. Or, use a CSL file, several of which are vendored in [_extensions/nature/csl](./_extensions/nature/csl/). The advantage of the latter is that references will be consistent across html, docx, and pdf outputs. See <https://pandoc.org/MANUAL.html#citation-syntax> for additional details on pandoc citation styles.

@@

# References
:::{#refs}
:::

@@

![Pyro-Velocity outperforms other RNA-velocity methods and accurately predicts cell lineage trajectories. a,b. RNA Velocity model comparisons from fully mature peripheral blood mononuclear cells (PBMCs, t-SNE embedded plots, a) or E15.5 mouse pancreas (UMAP embedded plots, b-c).  From left to right in panels a and b: 1. Cells colored by cell type; 2. Stream plot of velocity vector field defined by scVelo; 3. Stream plot of the posterior mean vector field from Pyro-Velocity; 4. Pyro-Velocity visualization of cell dynamics for 3 arbitrarily selected single cells; 5. Single-cell vector field uncertainty where each cell is colored based on the scaled angular standard deviation to visualize uncertainty based on n=30 iterated, posterior samples depicted as arrows noted throughout our analysis; 6. Averaged vector field uncertainty where each vector is colored by the averaged vector field uncertainty. The scaled angular standard deviation range is [0 360], where 0 corresponds to the highest confidence for a single direction, 360 corresponds to the highest uncertainty, and no directionality preference based on 30 posterior samples (4-6). Vectors were projected into the UMAP space using the scVelo API based on cosine similarity-derived transition matrix. c. UMAP embedded cells colored by the average cell shared time across 30 posterior samples. The Spearman’s correlation between Pyro-Velocity and Cytotrace is provided. d. Pyro-Velocity automatic gene selection and ranking for predicting important genes involved in pancreas development. e. Phase portraits of posterior samples average spliced and unspliced gene expression (left), rainbow plots showing denoised spliced gene expression (middle), and gene expression within UMAP embedded space.  Single dots correspond to individual cells (left, right) or vertical lines (middle). Cell type annotation coloring is the same as in panel b.](figures/fig.pdf){#fig-model-results}

![Pyro-Velocity accurately predicts cell fate choices and lineage trajectory in human LARRY barcoded hematopoietic cells. Systematic benchmarking of cell fate prediction methods - predicting uni-lineage cell fates (a-b, Model 1), bifurcated lineages (c, Model 2), or all lineages together (d, Model 2). Panels from left to right show on the same SPRING embedding: 1. Cell types; 2. Clonal progression vector field derived based on the observed clonal barcodes (Methods); 3. scVelo recovered vector field; 4. Pyro-Velocity single cell level vector field uncertainty (angular standard deviation from 30 posterior samples); 5. Pyro-Velocity recovered vector field; vectors are colored by the averaged vector field uncertainty; 6. Clonal fate potency score using Cospar [@Wang2022-xb]; 7. scVelo recovered latent time; 8. Pyro-Velocity inference of posterior mean shared time per cell. Cosine similarity and Spearman’s correlation for the different methods with the clonal progression and fate potency are reported in the titles.](figures/fig.pdf){#fig-lineage-tracing}

@@

![Graphical representations detail the generative process of the Pyro-Velocity models and highlight their differences. a. Pyro-Velocity Model 1. Each rectangle represents one data dimension (e.g., gene or cell). The arrows represent the directed hierarchical dependency among random variables. The white circles correspond to the modeled random variables. The probability distributions of these variables are denoted on the right legend. The gray circles correspond to intermediate results computed from random variables. In this model, the kinetics random variables for the cell shared time and gene are modeled with two constraints: 1. the shared time is strictly larger than the gene-specific activation time $t_{0}^{(0)}$, 2. the gene expression in the activate state starts with the initial condition $(u_{0}=0,\,s_{0}=0)$, as in previous methods[@La_Manno2018-lj;@Bergen2020-pj] (Methods). b. Pyro-Velocity Model 2. This figure follows the same graphical convention as in a. In this model $t_{0}^{(0)}$ is independent from the shared time, thus allowing a time lag for the gene activation phase. In addition, a positive starting basal gene expression level $(u_{0},\,s_{0})$ different than $(u_{0}=0,\,s_{0}=0)$ is allowed. (Methods). c. Schematic illustrating how the constraints proposed in Model 1 (left) and Model 2 (right) are reflected in the modeling of individual gene expression. The x-axis corresponds to the shared time, the y-axis corresponds to the posterior mean of the spliced expression level, the vertical lines with arrows show the start and end time of the observed cell sampling, t0 represents the gene-specific activation time $t_{0}^{(0)}$, the red lines highlight the differences for the potential gene-specific time lag when $t_{0}^{(0)} > {\min}\!⁡(t)$, and the initial spliced gene expression. The same differences can also be illustrated for the unspliced gene expression profiles (not shown).](figures/fig.pdf){#suppfig-graphical-models}

![Extended version of @fig-model-results for evaluating Pyro-Velocity (Model 1) and scVelo on the PBMC and pancreas datasets. a. Left figure shows the scVelo vector field using the default user options, right figure shows the evaluation of the scVelo vector field using different combinations of user options within Epsilon cells. b. Pyro-Velocity additional single cell level uncertainty diagnostic plots on the fully mature PBMC dataset, from left to right: 1. standard deviation of posterior samples of shared time, 2. Rayleigh test of posterior samples of the vector field (the title reports false discovery rate after p-value correction for multiple tests across cells), 3. transcriptional state uncertainty of predicted raw read count per cell. The contour lines show the top 10% of the most uncertain cells. c. For the pancreas dataset, the figures show from left to right: UMAP rendering of 1. Cytotrace, 2. scVelo latent time, 3. Posterior mean of Pyro-Velocity shared time averaged across 30 posterior samples, 4. Top 50 markers overlapping between Pyro-Velocity and scVelo (Methods, Supplementary Table 1-2), 5. The same uncertainty diagnostic plots per cell as in b. d. The same plot as Fig. 2d. e. Selection of the top negatively correlated genes with the cell shared time based on the posterior mean on denoised spliced expression and negative mean absolute error of gene expression prediction (Methods). f. Phase portraits, rainbow plots, and UMAP rendering of posterior mean of gene expression for the genes selected in e.](figures/fig.pdf){#suppfig-model-results}

![Extension of @fig-model-results, Pyro-Velocity (Model 2) outperforms scVelo on PBMC and pancreas datasets. Same format as in @fig-model-results and @suppfig-model-results](figures/fig.pdf){#suppfig-model-2-result}

![Extension of @fig-lineage-tracing, Pyro-velocity (Model 1 for uni-fate data, Model 2 for bi-fate and multi-fate data) outperforms Cytotrace on scRNA-seq datasets with lineage barcoding information (LARRY dataset). a. SPRING embedding rendering of Cytotrace scores, the title reports the Spearman’s correlation between the Cytotrace and Cospar fate potency scores, b. Pyro-Velocity additional single cell level uncertainty diagnostic plots based on the multi-fate LARRY dataset, from left to right: 1. Standard deviation of posterior samples of shared time per cell, 2. Rayleigh test of posterior samples of vector field per cell (the title reports the false discovery rate after p-value correction for multiple tests across cells), 3. Transcriptional state uncertainty of predicted raw read count per cell. The contour lines show the top 10% of the most uncertain cells (Methods), c. Pairwise Spearman’s correlation between Cytotrace/Cospar fate potency scores and velocity methods for the four different benchmarking datasets selected based on the lineage barcoding information from LARRY data. d. SPRING embedding rendering of posterior mean of shared time across 30 posterior samples, the title shows the Spearman’s correlation of mean shared time with Cytotrace. e. Selection of the top positively correlated genes with the cell shared time based on the posterior mean of spliced expression. f. Phase portraits, rainbow plots, and UMAP rendering with splicing gene expression level for the selected genes in e.](figures/fig.pdf){#suppfig-lineage-tracing}

![Extension of @fig-lineage-tracing presenting an extended comparison of Pyro-Velocity and scVelo on scRNA-seq datasets with lineage barcoding information (LARRY dataset). a. uni-fate Monocyte lineage analysis with Pyro-Velocity Model 2, b. uni-fate Neutrophil lineage analysis with Pyro-Velocity Model 2, c. Neutrophil and Monocyte bi-fate lineage analysis with Pyro-Velocity Model 1, and d. multi-fate lineage analysis with Pyro-Velocity Model 1. The panel descriptions are the same as in @fig-lineage-tracing](figures/fig.pdf){#suppfig-lineage-tracing-2}

>>>>

---
title: |
  Pyro&thinsp;-Velocity: Probabilistic RNA Velocity inference from single-cell data
number-sections: true
toc: true
toc-title: Contents
format:
  nature-pdf:
    # journal.cite-style is included in the tex file but ignored by pandoc if 
    # cite-method is not `natbib`.
    journal:
      cite-style: sn-nature
    classoption: 
      - lineno
    # `citeproc` is the pandoc default. Set `cite-method: natbib` if required
    # to use the bst styles from the upstream template.
    cite-method: natbib
    keep-tex: true
    equal-margins: false
    include-in-header: header.tex
    link-citations: true
  html:
    code-fold: true
    html-math-method: mathjax
  docx:
    reference-doc: styles.docx
# filters:
#   - authors-block
csl: _extensions/sciexp/nature/csl/springer-basic-brackets-no-et-al.csl
affiliations:
  - name: Massachusetts General Hospital Research Institute, Charlestown, MA
    id: 1
    department: Molecular Pathology Unit
    city: Charlestown
    state: MA
    postal-code: 02129
  - name: Harvard Medical School, Charlestown, MA
    id: 2
    department: Massachusetts General Hospital Cancer Center
    city: Charlestown
    state: MA
    postal-code: 02129
  - name: Massachusetts General Hospital, Boston, MA
    id: 3
    department: Center for Regenerative Medicine
    city: Boston
    state: MA
    postal-code: 02114
  - name: Harvard University, Cambridge, MA
    id: 4
    department: Harvard Stem Cell Institute
    city: Cambridge
    state: MA
    postal-code: 02139
  - name: Cambridge, MA
    id: 5
    department: Broad Institute of MIT and Harvard
    city: Cambridge
    state: MA
    postal-code: 02142
  - name: Laboratory of Neurodevelopmental Systems Biology, Lausanne, Switzerland
    id: 6
    department: |
      Brain and Mind Institute, School of Life Sciences, 
      École Polytechnique Fédérale de Lausanne (EPFL)
    city: Lausanne
    state: Switzerland
    postal-code: 1015
  - name: Cambridge, MA
    id: 7
    department: Basis Research Institute
    city: Cambridge
    state: MA
    postal-code: 02142
authors:
  - name: Qian Qin
    affiliations:
      - ref: 1
      - ref: 2
      - ref: 3
      - ref: 4
      - ref: 5
    # attributes:
    #   equal-contributor: true
  - name: Eli Bingham
    affiliations:
      - ref: 5
      - ref: 7
  - name: Gioele La Manno
    affiliations:
      - ref: 6
    email: gioele.lamanno@epfl.ch
    attributes:
      corresponding: true
  - name: David M. Langenau
    affiliations:
      - ref: 1
      - ref: 2
      - ref: 3
      - ref: 4
    email: dlangenau@mgh.harvard.edu
    attributes:
      corresponding: true
  - name: Luca Pinello
    affiliations:
      - ref: 1
      - ref: 2
      - ref: 5
    email: lpinello@mgh.harvard.edu
    attributes:
      corresponding: true
abstract: |
  Single-cell RNA Velocity has dramatically advanced our ability to model
  cellular differentiation and cell fate decisions. However, current
  preprocessing choices and model assumptions often lead to errors in assigning
  developmental trajectories. Here, we develop, Pyro-Velocity, a Bayesian,
  generative, and multivariate RNA Velocity model to estimate the uncertainty of
  cell future states. This approach models raw sequencing counts with the
  synchronized cell time across all expressed genes to provide quantifiable and
  improved information on cell fate choices and developmental trajectory
  dynamics.
keywords: [
  RNA velocity, 
  single-cell RNA sequencing, 
  probabilistic generative modeling, 
  cell fate, 
  developmental trajectory inference
]
bibliography: bibliography.bib
crossref:
  fig-prefix: Fig.
  custom:
    - kind: float
      key: suppfig
      latex-env: suppfig
      reference-prefix: Fig. S
      space-before-numbering: false
      latex-list-of-description: Supplementary Figure
---

@@

@@

# References

:::{#refs}
:::

@@

::: {#fig-model-results}
![](figures/fig.pdf)

Pyro-Velocity outperforms other RNA-velocity methods and accurately predicts
cell lineage trajectories. a,b. RNA Velocity model comparisons from fully mature
peripheral blood mononuclear cells (PBMCs, t-SNE embedded plots, a) or E15.5
mouse pancreas (UMAP embedded plots, b-c).  From left to right in panels a and
b: 1. Cells colored by cell type; 2. Stream plot of velocity vector field
defined by scVelo; 3. Stream plot of the posterior mean vector field from
Pyro-Velocity; 4. Pyro-Velocity visualization of cell dynamics for 3 arbitrarily
selected single cells; 5. Single-cell vector field uncertainty where each cell
is colored based on the scaled angular standard deviation to visualize
uncertainty based on n=30 iterated, posterior samples depicted as arrows noted
throughout our analysis; 6. Averaged vector field uncertainty where each vector
is colored by the averaged vector field uncertainty. The scaled angular standard
deviation range is [0 360], where 0 corresponds to the highest confidence for a
single direction, 360 corresponds to the highest uncertainty, and no
directionality preference based on 30 posterior samples (4-6). Vectors were
projected into the UMAP space using the scVelo API based on cosine
similarity-derived transition matrix. c. UMAP embedded cells colored by the
average cell shared time across 30 posterior samples. The Spearman’s correlation
between Pyro-Velocity and Cytotrace is provided. d. Pyro-Velocity automatic gene
selection and ranking for predicting important genes involved in pancreas
development. e. Phase portraits of posterior samples average spliced and
unspliced gene expression (left), rainbow plots showing denoised spliced gene
expression (middle), and gene expression within UMAP embedded space.  Single
dots correspond to individual cells (left, right) or vertical lines (middle).
Cell type annotation coloring is the same as in panel b.
:::

::: {#fig-lineage-tracing}
![](figures/fig.pdf)

Pyro-Velocity accurately predicts cell fate choices and lineage trajectory in
human LARRY barcoded hematopoietic cells. Systematic benchmarking of cell fate
prediction methods - predicting uni-lineage cell fates (a-b, Model 1),
bifurcated lineages (c, Model 2), or all lineages together (d, Model 2). Panels
from left to right show on the same SPRING embedding: 1. Cell types; 2. Clonal
progression vector field derived based on the observed clonal barcodes
(Methods); 3. scVelo recovered vector field; 4. Pyro-Velocity single cell level
vector field uncertainty (angular standard deviation from 30 posterior samples);
5. Pyro-Velocity recovered vector field; vectors are colored by the averaged
vector field uncertainty; 6. Clonal fate potency score using Cospar
[@Wang2022-xb]; 7. scVelo recovered latent time; 8. Pyro-Velocity inference of
posterior mean shared time per cell. Cosine similarity and Spearman’s
correlation for the different methods with the clonal progression and fate
potency are reported in the titles.
:::

@@

::: {#suppfig-graphical-models}
![](figures/fig.pdf)

Graphical representations detail the generative process of the Pyro-Velocity
models and highlight their differences. a. Pyro-Velocity Model 1. Each rectangle
represents one data dimension (e.g., gene or cell). The arrows represent the
directed hierarchical dependency among random variables. The white circles
correspond to the modeled random variables. The probability distributions of
these variables are denoted on the right legend. The gray circles correspond to
intermediate results computed from random variables. In this model, the kinetics
random variables for the cell shared time and gene are modeled with two
constraints: 1. the shared time is strictly larger than the gene-specific
activation time $t_{0}^{(0)}$, 2. the gene expression in the activate state
starts with the initial condition $(u_{0}=0,\,s_{0}=0)$, as in previous
methods[@La_Manno2018-lj;@Bergen2020-pj] (Methods). b. Pyro-Velocity Model 2.
This figure follows the same graphical convention as in a. In this model
$t_{0}^{(0)}$ is independent from the shared time, thus allowing a time lag for
the gene activation phase. In addition, a positive starting basal gene
expression level $(u_{0},\,s_{0})$ different than $(u_{0}=0,\,s_{0}=0)$ is
allowed. (Methods). c. Schematic illustrating how the constraints proposed in
Model 1 (left) and Model 2 (right) are reflected in the modeling of individual
gene expression. The x-axis corresponds to the shared time, the y-axis
corresponds to the posterior mean of the spliced expression level, the vertical
lines with arrows show the start and end time of the observed cell sampling, t0
represents the gene-specific activation time $t_{0}^{(0)}$, the red lines
highlight the differences for the potential gene-specific time lag when
$t_{0}^{(0)} > {\min}\!⁡(t)$, and the initial spliced gene expression. The same
differences can also be illustrated for the unspliced gene expression profiles
(not shown).
:::

::: {#suppfig-model-results}
![](figures/fig.pdf)

Extended version of @fig-model-results for evaluating Pyro-Velocity (Model 1)
and scVelo on the PBMC and pancreas datasets. a. Left figure shows the scVelo
vector field using the default user options, right figure shows the evaluation
of the scVelo vector field using different combinations of user options within
Epsilon cells. b. Pyro-Velocity additional single cell level uncertainty
diagnostic plots on the fully mature PBMC dataset, from left to right: 1.
standard deviation of posterior samples of shared time, 2. Rayleigh test of
posterior samples of the vector field (the title reports false discovery rate
after p-value correction for multiple tests across cells), 3. transcriptional
state uncertainty of predicted raw read count per cell. The contour lines show
the top 10% of the most uncertain cells. c. For the pancreas dataset, the
figures show from left to right: UMAP rendering of 1. Cytotrace, 2. scVelo
latent time, 3. Posterior mean of Pyro-Velocity shared time averaged across 30
posterior samples, 4. Top 50 markers overlapping between Pyro-Velocity and
scVelo (Methods, Supplementary Table 1-2), 5. The same uncertainty diagnostic
plots per cell as in b. d. The same plot as Fig. 2d. e. Selection of the top
negatively correlated genes with the cell shared time based on the posterior
mean on denoised spliced expression and negative mean absolute error of gene
expression prediction (Methods). f. Phase portraits, rainbow plots, and UMAP
rendering of posterior mean of gene expression for the genes selected in e.
:::

::: {#suppfig-model-2-result}
![](figures/fig.pdf)

Extension of @fig-model-results, Pyro-Velocity (Model 2) outperforms scVelo on
PBMC and pancreas datasets. Same format as in @fig-model-results and
@suppfig-model-results
:::

::: {#suppfig-lineage-tracing}
![](figures/fig.pdf)

Extension of @fig-lineage-tracing, Pyro-velocity (Model 1 for uni-fate data,
Model 2 for bi-fate and multi-fate data) outperforms Cytotrace on scRNA-seq
datasets with lineage barcoding information (LARRY dataset). a. SPRING embedding
rendering of Cytotrace scores, the title reports the Spearman’s correlation
between the Cytotrace and Cospar fate potency scores, b. Pyro-Velocity
additional single cell level uncertainty diagnostic plots based on the
multi-fate LARRY dataset, from left to right: 1. Standard deviation of posterior
samples of shared time per cell, 2. Rayleigh test of posterior samples of vector
field per cell (the title reports the false discovery rate after p-value
correction for multiple tests across cells), 3. Transcriptional state
uncertainty of predicted raw read count per cell. The contour lines show the top
10% of the most uncertain cells (Methods), c. Pairwise Spearman’s correlation
between Cytotrace/Cospar fate potency scores and velocity methods for the four
different benchmarking datasets selected based on the lineage barcoding
information from LARRY data. d. SPRING embedding rendering of posterior mean of
shared time across 30 posterior samples, the title shows the Spearman’s
correlation of mean shared time with Cytotrace. e. Selection of the top
positively correlated genes with the cell shared time based on the posterior
mean of spliced expression. f. Phase portraits, rainbow plots, and UMAP
rendering with splicing gene expression level for the selected genes in e.
:::

::: {#suppfig-lineage-tracing-2}
![](figures/fig.pdf)

Extension of @fig-lineage-tracing presenting an extended comparison of
Pyro-Velocity and scVelo on scRNA-seq datasets with lineage barcoding
information (LARRY dataset). a. uni-fate Monocyte lineage analysis with
Pyro-Velocity Model 2, b. uni-fate Neutrophil lineage analysis with
Pyro-Velocity Model 2, c. Neutrophil and Monocyte bi-fate lineage analysis with
Pyro-Velocity Model 1, and d. multi-fate lineage analysis with Pyro-Velocity
Model 1. The panel descriptions are the same as in @fig-lineage-tracing
:::
